---
title: "Stock price prediction with time series techniques"
author: "Yevheniia Mitriakhina"
output:
  html_document:
    df_print: paged
---
# Getting the data
```{r}

library(BatchGetSymbols)

first.date <- '2008-01-02'
last.date <- Sys.Date()-30
freq.data <- 'daily'
# set tickers
tickers <- "HSY"

l.out <- BatchGetSymbols(tickers = tickers, 
                         first.date = first.date,
                         last.date = last.date, 
                         freq.data = freq.data,
                         cache.folder = file.path(tempdir(), 
                                                  'BGS_Cache'),
                         thresh.bad.data = 0.99)
test.data <- BatchGetSymbols(tickers = tickers, 
                         first.date = last.date,
                         last.date = Sys.Date(), 
                         freq.data = freq.data,
                         cache.folder = file.path(tempdir(), 
                                                  'BGS_Cache'),
                         thresh.bad.data = 0.99)
```
```{r}
data.prices <- ts(l.out$df.tickers$price.close, start=c(2008, 1,2), frequency = 251)
test.set <- ts(test.data$df.tickers$price.close, start=c(2019, 4, 4), frequency = 251)
```
# Preparing the data
we need to understand the three components of a time series data:
Trend,Seasonal, Cyclic.
Smoothing with moving average
```{r}
#Simple moving average
library(TTR)
smoothed <- SMA(data.prices, n=75)
plot(ts(smoothed, start=c(2008, 1,2), frequency = 251), col='red', main="Simple moving average on prices", ylab='Price')
lines(data.prices, col='lightgrey')
legend(2008, 100, legend=c("Price", "Moving average"),
       col=c("grey", "red"), lty=1:1, cex=0.8)
```

# Stationarity test
Augmented Dickey-Fuller test (ADF)

ADF tests the null hypothesis that a unit root is present in time series sample. ADF statistic is a negative number and more negative it is the stronger the rejection of the hypothesis that there is a unit root.
Null Hypotehsis (H0): If accepted, it suggests the time series has a unit root, meaning it is non-stationary. It has some time dependent structure.
```{r}
library(tseries)
tseries::adf.test(data.prices)
```
we cannot reject the hypothesis that data is non-stationary on 95% confidence level.
```{r}
log.prices <- log(data.prices)
plot(log.prices, main='Prices', ylab='log(price)', col='grey')
```

```{r}
decomposed <- stl(log(data.prices), s.window="periodic")
plot(decomposed, col = '#708090', main = "Decomposition of additive time series") 
```
We have increasing trend which is almost linear we might fit the line to the data and then model the residuals from that line.
```{r}
library(forecast)
x <- seq(1, length(log.prices), by=1)
plot(cbind(x, as.vector(log.prices)), type='l', col='grey', main='Fitting the linear trend', ylab='log(Price)', xlab='Time')
fit <- tslm(log.prices ~ trend)
intercept <- data.frame(fit$coefficients)[1,]
slope <- data.frame(fit$coefficients)[2,]
y <- intercept+slope*x
lines(cbind(x, y), col='red')
legend(2008, 4, legend=c("Price", "Trend"),
       col=c("grey", "red"), lty=1:1, cex=0.8)
```
```{r}
detrended.prices <- log.prices - y
plot(detrended.prices, ylab='Price', main='Detrended prices', col='grey')
abline(h = mean(detrended.prices), lty = 2, col='red')
legend(2008, 0.3, legend=c("Detrended log(Price)", "Mean"),
       col=c("grey", "red"), lty=1:1, cex=0.8)
```
Even visually the data is still not stationary. Again check for stationarity:
```{r}
# Time series has non-zero mean
library(tseries)
tseries::adf.test(detrended.prices) # Still non-stationary
```
# Difference detrended data
```{r}
differenced.prices <- diff(detrended.prices) 
plot(differenced.prices, main='Differenced prices (d=1)', ylab='log(Price)', col='grey')
abline(h = mean(differenced.prices), lty = 2, col='red')
legend(2008, 0.15, legend=c("Differenced detrended log(Price)", "Mean"),
       col=c("grey", "red"), lty=1:1, cex=0.8)

```
```{r}
tseries::adf.test(differenced.prices)
```
```{r}
# Examining 
Acf(differenced.prices, main='ACF for Differenced Series', lag.max=250)
Pacf(differenced.prices, main='PACF for Differenced Series', lag.max = 250)
```

```{r}
# Automatically determines coefficients
 auto.arima(differenced.prices)
```

```{r}
# Choosing the best fit
f1 <- arima(differenced.prices, c(2,1,0))
f2 <- arima(differenced.prices, c(2,1,2))
f3 <- arima(differenced.prices, c(2,1,1))
f4 <- arima(differenced.prices, c(0,0,2))
# The best fit
fit <- arima(differenced.prices, c(2,0, 0))
fit
```

```{r}
fcast <- forecast(fit, h=30)
fcast
```
```{r}
# Perform tha same transformations on test set to compare it with prediction later
test.log <- log(test.set)
test.detrended <- test.log - intercept+slope*seq((length(log.prices) + 1), (length(log.prices) + length(test.log)), by=1)
test.differenced <- diff(test.detrended)

```

```{r}
plot(as.numeric(test.differenced), col = 'grey', type = 'l', main = 'Predicted and real prices', xlab='Time', ylab = 'Price')
lines(as.numeric(fcast$x[length(fcast$x) - 30 : length(fcast$x)]), col='red')
legend(1, 0.04, legend=c("Differenced detrended log(price)", "Prediction"),
       col=c("grey", "red"), lty=1:1, cex=0.8)

```


